---
title: "a3_task3_Drew_Andersen"
author: "Drew Andersen"
date: "10/24/2021"
output: html_document
---
# Overview
This code wrangles and visualizes data for red and purple urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers.

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) #Attach packages
library(janitor)
library(here)
library(lubridate)
```

# Section 1
This section will compare the size of purple and red urchins.

# Read in the data
```{r}
urchins <- read_csv(here("data","urchins.csv")) %>% #read in data
  clean_names() #convert to lower snake case
```

#Data Visualization
Creating a faceted histogram to compare the distribution of sizes between purple and red urchins.
```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) +
  facet_wrap(vars(common_name)) +
  labs(title = "Size Distribution of Purple & Red Urchins",
       x = "Urchin Size (cm)",
       y = "Count",
       caption = "Distribution of urchin sizes according to their color. Size is recorded by measuring the urchin's diameter size to the nearest 0.5 cm") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("purple","red"))
```

