---
title: "Santa Barbara Urchins"
author: "Drew Andersen"
date: "10/24/2021"
output: html_document
---
# Overview
This code wrangles and visualizes data for red and purple urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers.

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

# Setup
```{r setup, include = TRUE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse) #Attach packages
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

# Read in the data
```{r}
urchins <- read_csv(here("data","urchins.csv")) %>% #read in data
  clean_names() #convert to lower snake case
```

# {.tabset .tabset-fade}

## Section 1
This section will compare the size of purple and red urchins.

### Data Wrangling 
Calculating summary statistics for each type of urchin.
```{r}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(sample_size = n(), #summary statistics for both species of urchin
            mean_size = round(mean(size),1),
            std_dev = round(sd(size),1),
            median_size = median(size))

urchins_stats %>% 
  kable(col.names = c("Species", #renaming the column names from their variable names
                      "Sample Size", 
                      "Mean Diameter (cm)",
                      "Standard Deviation (cm)",
                      "Median Diameter (cm)")) %>% 
  kable_styling(full_width = FALSE)
```

### Data Visualization
Creating a faceted histogram to compare the distribution of sizes between purple and red urchins.

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) + #histogram that is colored according to the common_name
  scale_fill_manual(values = c("purple","red")) + #specific colors chosen
  facet_wrap(vars(common_name)) + #creates two seperate side by side histograms
  labs(title = "Size Distribution of Purple & Red Urchins", #adds labels/titles
       x = "Urchin Diameter (cm)",
       y = "Count",
       caption = "Figure 1. Distribution of urchin sizes by species. Size is recorded by measuring the urchin's diameter to the nearest 0.5 cm") +
  theme_bw() +
  theme(legend.position = "none") #removes redundant legend 
```

Comparing the two plots above, we can see that purple urchins are larger than red urchins on average. Purple urchins also appear more narrow in their distribution than red urchins. The distribution of purple urchins is also positively skewed, while the red urchins appear more normally distributed. Using the values from the table above, we can calcualte the percent difference in size between the two urchin species to be `r round((urchins_stats$mean_size[2] - urchins_stats$mean_size[1])*100/urchins_stats$mean_size[1],0)`%.

## Section 2
This section will examine the number of purple urchins observed at various sites by year.

### Data Wrangling
Creates a subset of data for purple urchin counts at various reefs between 2008-2020.
```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% #discludes the red urchin population
  mutate(date_new = mdy(date)) %>% #parses the date from this column 
  mutate(year = year(date_new)) %>% #creates a column for year that can now be grouped by 
  mutate(site_full = case_when( #replaces acronym with full reef name
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef")) %>% 
  group_by(site_full, year) %>% 
  summarize(count = n()) #finds the count of purple urchins by site by year
```

### Data Visualization 
Creates a line plot where each series represents a different sample site. 
```{r}
ggplot(data = purple_urchins, aes(x = year, y = count)) +
  geom_line(aes(color = site_full)) + #creates a series for each site
  labs(title = "Purple Urchin Count by Sample Site",
       y = "Total Count",
       color = "Sample Site",
       caption = "Fig 2. Purple Urchin Count by Reef Site between 2008-2020. Data collection for Isla Vista Reef did not begin until 2012.") +
  scale_x_continuous(breaks = c(2008,2010, 2012, 2014, 2016, 2018, 2020)) + #causes breaks on the x-axis at these values
  theme_linedraw() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = c(.75,.75),
         plot.caption = element_text(hjust = 0))
```

No, I don't believe this graph reflects the actual urchin populations in Santa Barbara. In order to determine whether this graph is reflective of these populations, we would have to take into account the number of dives per year at each site, as well as the time and length of these dives. Other inconsistencies in sampling practices could affect the total count of pruple urchins as well.

