---
title: "a3_task3_Drew_Andersen"
author: "Drew Andersen"
date: "10/24/2021"
output: html_document
---
## Overview
This code wrangles and visualizes data for red and purple urchins recorded by Santa Barbara Coastal Long Term Ecological Research program divers.

Data Citation: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Setup
```{r setup, include = TRUE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse) #Attach packages
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

# Section 1
This section will compare the size of purple and red urchins.

## Read in the data
```{r}
urchins <- read_csv(here("data","urchins.csv")) %>% #read in data
  clean_names() #convert to lower snake case
```

## Data Wrangling 
Calculating summary statistic for each type of urchin.
```{r}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(sample_size = n(), #summary statistics for both species of urchin
            mean_size = mean(size),
            std_dev = sd(size),
            median_size = median(size))

urchins_stats %>% 
  kable(col.names = c("Species", #renaming the column names from their variable names
                      "Sample Size", 
                      "Mean Diameter (cm)",
                      "Standard Deviation (cm)",
                      "Median Diameter (cm)")) %>% 
  kable_styling(full_width = FALSE)
```

## Data Visualization
Creating a faceted histogram to compare the distribution of sizes between purple and red urchins.

```{r}
ggplot(data = urchins, aes(x = size)) +
  geom_histogram(bins = 10, aes(fill = common_name)) + #histogram that is colored according to the common_name
  scale_fill_manual(values = c("purple","red")) + #specific colors chosen
  facet_wrap(vars(common_name)) + #creates two seperate side by side histograms
  labs(title = "Size Distribution of Purple & Red Urchins", #adds labels/titles
       x = "Urchin Diameter (cm)",
       y = "Count",
       caption = "Figure 1. Distribution of urchin sizes by species. Size is recorded by measuring the urchin's diameter to the nearest 0.5 cm") +
  theme(legend.position = "none") #removes redundant legend 
```
Comparing the two plots above, we can see that red urchins are larger than purple urchins by roughly 50%. The distribution of size for purple urchins is positively skewed and has a narrower spread as well. The red urchins are larger and also have a broader distribution in size. 

# Section 2
This section will examine the number of purple urchins observed at various sites by year.

## Data Wrangling
```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef")) %>% 
  group_by(site_full, year) %>% 
  summarize(count = n())
```

## Data Visualization 
```{r}
ggplot(data = purple_urchins, aes(x = year, y = count)) +
  geom_line(aes(color = site_full)) +
  labs(title = "Purple Urchin Count by Sample Site",
       y = "Total Count",
       x = "Year",
       color = "Sample Site",
       caption = "Fig 2. Purple Urchin Count by Reef Site between 2008-2020") +
  scale_x_continuous(breaks = c(2008,2010, 2012, 2014, 2016, 2018, 2020))
```
No, I don't believe this graph reflects the actual urchin populations in Santa Barbara. In order to determine whether this graph is reflective of these populations, we would have to take into account the number of dives per year at each site, as well as the time and length of these dives. Other inconsistencies in sampling practices could affect the total count of pruple urchins as well.

